Final Project Paper

### 2017-03-08 paper/paper.Rmd
updated intro

Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas. The Federal Bureau of Investigation (FBI) collects crime data every year for individual cities within the United States, and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010. There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis does not consider outside covariates when comparing crime rate and GDP of metropolitan areas. Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account. One other research group took into account economic uncertainty whereas we will not be taking into account outside covariates and comparing crime rate and GDP at face value. We want to look at the GDP and crime relationship unaffected by other covariates to see if there is a correlation between the two variables.

Our analysis focuses on violent and property crime impact on city GDPs without covariates. The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the city level, adding an extra area of granularity. In terms of covariates, we are not accounting for confounding covariates such as uncertainty demonstrated by Goulas. We are observing and comparing violent and property crime trends to city GDP trends.

>>>>

Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas. The Federal Bureau of Investigation (FBI) collects crime data and population every year for individual cities within the United States and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010. There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis considers population as another independent variable when comparing crime rate and GDP of metropolitan areas. Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account. One other research group took into account economic uncertainty whereas we will focus more on the count data of a metropolitan area (population, GDP). We want to look at the GDP and crime relationship affected by population to see if GDP per person and population of a metropolitan area have any sort of impact on the crimes per person committed within that metropolitan area.

Our analysis focuses on violent and property crime impact on city GDPs with population as another independent variable. The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the city level, adding an extra area of granularity. In terms of covariates, we are accounting for the population change between years along with the changes in GDP and crime. We are observing and comparing violent and property crime trends to metropolitan area GDP and population trends.
